#Practice Machine Learning - Project
##author: "Chun Lei He"
##date: "June 8, 2016"
##output: html_document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("kernlab")) install.packages("kernlab")
if (!require("RANN")) install.packages("RANN")

library(caret)
library(ggplot2)
library(lattice)
library(kernlab)
library(randomForest)
```

##Background

###Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of 
###data about personal activity relatively inexpensively. These type of devices are part of the quantified self 
###movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, 
###to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is 
###quantify how much of a particular activity they do, but they rarely quantify how well they do it. 
###In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 
###6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 
###More information is available from the website here: http://groupware.les.inf.puc-rio.br/har 
###(see the section on the Weight Lifting Exercise Dataset).

##Data

###The training data for this project are available here:

###  https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

###The test data are available here:

###  https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

##Load Data 

```{r}
setInternet2(TRUE)
download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile = "pml-training.csv")
download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile = "pml-testing.csv")
pmltrain <- read.csv('pml-training.csv')
pmltest <- read.csv('pml-testing.csv')

```

##Exploratory Data Analysis


###Split training, test and validation sets
```{r}
trainidx <- createDataPartition(pmltrain$classe,p=.9,list=FALSE)
traindata = pmltrain[trainidx,]
validationdata = pmltrain[-trainidx,]
set.seed(123456)

```

###Clean up near zero variance features, columns with missing values and descriptive fields.

```{r}
nzv <- nearZeroVar(traindata)
trainnzv <- traindata[-nzv]
validationnzv <- validationdata[-nzv]
pmltestnzv <- pmltest[-nzv]


ftridx <- which(lapply(trainnzv,class) %in% c('numeric'))
trainnzv1 <- preProcess(trainnzv[,ftridx], method=c('knnImpute'))

ftridx
trainnzv1

```

###Created from 367 samples and 64 variables

###Pre-processing:
###- centered (64)
###- ignored (0)
###- 5 nearest neighbor imputation (64)
###- scaled (64)

```{r}
pred1 <- predict(trainnzv1, trainnzv[,ftridx])
predtrain <- cbind(trainnzv$classe,pred1)
names(predtrain)[1] <- 'classe'
predtrain[is.na(predtrain)] <- 0

pred2 <- predict(trainnzv1, validationnzv[,ftridx])
predvalidation <- cbind(validationnzv$classe, pred2)
names(predvalidation)[1] <- 'classe'
predvalidation[is.na(predvalidation)] <- 0

predpmltest <- predict(trainnzv1,pmltestnzv[,ftridx] )

```
## Build the random Forest model, train it on the training set

```{r}
model <- randomForest(classe~.,data=predtrain)

predtrain1 <- predict(model, predtrain) 
print(table(predtrain1, predtrain$classe))

training <- as.data.frame(table(predtrain1, predtrain$classe))

```

## Apply the model on the validation set
```{r}
predvalidation1 <- predict(model, predvalidation) 
print(table(predvalidation1, predvalidation$classe))

```
## Apply the model on test set
```{r}
str(predpmltest)
predanswers <- predict(model, predpmltest) 

```

## show the prediction results
```{r}
predanswers
```

