#Practice Machine Learning - Project
###Author: "Chun Lei He"
###Date: "June 8, 2016"
###Output: html_document


##Background

###Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of 
###data about personal activity relatively inexpensively. These type of devices are part of the quantified self 
###movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, 
###to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is 
###quantify how much of a particular activity they do, but they rarely quantify how well they do it. 
###In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 
###6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 
###More information is available from the website here: http://groupware.les.inf.puc-rio.br/har 
###(see the section on the Weight Lifting Exercise Dataset).

##Data

###The training data for this project are available here:

###  https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

###The test data are available here:

###  https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

### Install the packages and call the libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("kernlab")) install.packages("kernlab")
if (!require("RANN")) install.packages("RANN")

library(caret)
library(ggplot2)
library(lattice)
library(kernlab)
library(randomForest)
```


##Load Data 

```{r}
setInternet2(TRUE)
download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile = "pml-training.csv")
download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile = "pml-testing.csv")
pmltrain <- read.csv('pml-training.csv')
pmltest <- read.csv('pml-testing.csv')

```

##Exploratory Data Analysis


###Split training, test and validation sets
```{r}
trainidx <- createDataPartition(pmltrain$classe,p=.9,list=FALSE)
traindata = pmltrain[trainidx,]
validationdata = pmltrain[-trainidx,]
set.seed(123456)

```

###Clean up near zero variance features, columns with missing values and descriptive fields.

```{r}
nzv <- nearZeroVar(traindata)
trainnzv <- traindata[-nzv]
validationnzv <- validationdata[-nzv]
pmltestnzv <- pmltest[-nzv]


ftridx <- which(lapply(trainnzv,class) %in% c('numeric'))
trainnzv1 <- preProcess(trainnzv[,ftridx], method=c('knnImpute'))

ftridx
##[1]   7   8   9  11  13  15  17  18  19  20  21  22  23  24  25  26  27  28  29  36  37  38  40  41  42  43  50  52  54
##[30]  56  57  58  59  60  61  62  63  64  66  67  68  69  70  71  72  73  74  75  76  77  78  84  85  86  87  88  89  90
##[59]  91  93  94  95  96 101 102

trainnzv1
##Created from 369 samples and 65 variables
##
##Pre-processing:
##  - centered (65)
##  - ignored (0)
##  - 5 nearest neighbor imputation (65)
##  - scaled (65)
```

### Build the training, validation, and test sets with selected variables with class labels
```{r}
pred1 <- predict(trainnzv1, trainnzv[,ftridx])
predtrain <- cbind(trainnzv$classe,pred1)
names(predtrain)[1] <- 'classe'
predtrain[is.na(predtrain)] <- 0

pred2 <- predict(trainnzv1, validationnzv[,ftridx])
predvalidation <- cbind(validationnzv$classe, pred2)
names(predvalidation)[1] <- 'classe'
predvalidation[is.na(predvalidation)] <- 0

predpmltest <- predict(trainnzv1,pmltestnzv[,ftridx] )

```
## Build the random Forest model, train it on the training set

```{r}
model <- randomForest(classe~.,data=predtrain)

predtrain1 <- predict(model, predtrain) 
print(table(predtrain1, predtrain$classe))

##predtrain1    A    B    C    D    E
##         A 5022    0    0    0    0
##         B    0 3418    0    0    0
##         C    0    0 3080    0    0
##         D    0    0    0 2895    0
##         E    0    0    0    0 3247
         
training <- as.data.frame(table(predtrain1, predtrain$classe))

```

## Apply the model on the validation set
```{r}
predvalidation1 <- predict(model, predvalidation) 
print(table(predvalidation1, predvalidation$classe))

##predvalidation1   A   B   C   D   E
##              A 554   3   0   1   0
##              B   4 373   2   0   0
##              C   0   3 339   3   0
##              D   0   0   1 317   1
##              E   0   0   0   0 359
```
## Apply the model on test set
```{r}
str(predpmltest)

##'data.frame':	20 obs. of  65 variables:
## $ roll_belt               : num  0.933 -1.012 -1.014 0.964 -1.007 ...
## $ pitch_belt              : num  1.1923 0.2013 0.0647 -1.8798 0.1323 ...
## $ yaw_belt                : num  0.0686 -0.8169 -0.8127 1.8231 -0.8137 ...
## $ max_roll_belt           : num  0.049 -0.857 -0.855 1.836 -0.871 ...
## $ min_roll_belt           : num  0.0757 -0.8275 -0.8251 1.8583 -0.8419 ...
## $ amplitude_roll_belt     : num  -0.09096 -0.13658 -0.13733 -0.00441 -0.13658 ...
## $ var_total_accel_belt    : num  -0.307 -0.307 -0.325 -0.343 -0.378 ...
## $ avg_roll_belt           : num  0.871 -1.063 -1.064 0.929 -1.055 ...
## $ stddev_roll_belt        : num  -0.396 -0.438 -0.429 -0.281 -0.463 ...
## $ var_roll_belt           : num  -0.318 -0.321 -0.321 -0.3 -0.323 ...
## $ avg_pitch_belt          : num  1.131 0.185 0.184 -1.903 0.179 ...
## $ stddev_pitch_belt       : num  -0.539 0.0592 -0.2557 -0.1297 -0.2682 ...
## $ var_pitch_belt          : num  -0.3967 -0.1749 -0.3033 -0.0348 -0.3253 ...
## $ avg_yaw_belt            : num  0.0664 -0.8454 -0.843 1.8616 -0.8596 ...
## $ stddev_yaw_belt         : num  -0.0873 -0.1188 -0.1206 -0.0261 -0.1204 ...
## $ var_yaw_belt            : num  -0.0679 -0.068 -0.068 -0.0662 -0.068 ...
## $ gyros_belt_x            : num  -2.378 -0.262 0.267 0.556 0.171 ...
## $ gyros_belt_y            : num  -0.76 -0.76 -0.249 0.901 -0.249 ...
## $ gyros_belt_z            : num  -1.367 0.254 0.669 -0.12 0.545 ...
## $ roll_arm                : num  0.316 -0.244 -0.244 -1.743 0.803 ...
## $ pitch_arm               : num  -0.757 0.151 0.151 1.949 0.242 ...
## $ yaw_arm                 : num  2.50653 0.00702 0.00702 -1.98697 1.43932 ...
## $ var_accel_arm           : num  0.7373 -0.501 -0.0652 -0.8136 -0.3748 ...
## $ gyros_arm_x             : num  -0.849 -0.6083 1.0311 0.0885 -1.0044 ...
## $ gyros_arm_y             : num  0.865 1.299 -1.292 -0.296 1.229 ...
## $ gyros_arm_z             : num  -0.814 -1.266 1.552 1.173 -1.464 ...
## $ max_picth_arm           : num  0.884 -0.507 -0.507 -0.996 0.67 ...
## $ min_roll_arm            : num  -0.507 0.768 0.768 1.277 -0.393 ...
## $ amplitude_pitch_arm     : num  0.3704 -1.0359 -1.0359 0.4202 0.0715 ...
## $ roll_dumbbell           : num  -0.59 0.441 0.478 0.279 -1.785 ...
## $ pitch_dumbbell          : num  0.971 -1.16 -1.097 -0.519 -1.153 ...
## $ yaw_dumbbell            : num  1.507 -0.938 -0.934 -1.275 -0.195 ...
## $ max_roll_dumbbell       : num  0.718 -1.156 -1.079 -0.815 -0.143 ...
## $ max_picth_dumbbell      : num  1.162 -1.084 -0.988 -1.364 0.325 ...
## $ min_roll_dumbbell       : num  1.54 -0.507 -0.539 0.186 0.174 ...
## $ min_pitch_dumbbell      : num  1.839 -0.722 -0.632 -1.043 0.4 ...
## $ amplitude_roll_dumbbell : num  -0.344 -0.691 -0.604 -0.83 -0.235 ...
## $ amplitude_pitch_dumbbell: num  -0.43139 -0.73378 -0.69782 -0.76919 0.00996 ...
## $ var_accel_dumbbell      : num  -0.24 -0.217 -0.237 -0.281 -0.259 ...
## $ avg_roll_dumbbell       : num  -0.383 0.51 0.635 0.299 -0.576 ...
## $ stddev_roll_dumbbell    : num  -0.345 -0.637 -0.61 -0.668 -0.219 ...
## $ var_roll_dumbbell       : num  -0.38 -0.436 -0.433 -0.44 -0.31 ...
## $ avg_pitch_dumbbell      : num  1.3143 -1.1511 -1.0965 -0.5575 -0.0907 ...
## $ stddev_pitch_dumbbell   : num  -0.377 -0.667 -0.653 -0.819 -0.174 ...
## $ var_pitch_dumbbell      : num  -0.419 -0.488 -0.487 -0.51 -0.242 ...
## $ avg_yaw_dumbbell        : num  1.606 -0.982 -0.904 -1.317 0.337 ...
## $ stddev_yaw_dumbbell     : num  -0.4621 -0.7054 -0.7017 -0.7198 -0.0397 ...
## $ var_yaw_dumbbell        : num  -0.407 -0.451 -0.449 -0.45 -0.21 ...
## $ gyros_dumbbell_x        : num  1.205 0.434 0.563 -0.182 0.306 ...
## $ gyros_dumbbell_y        : num  0.033 0.0124 0.1977 -0.1316 -1.0579 ...
## $ gyros_dumbbell_z        : num  -1.454 -1.767 -0.61 0.609 -0.985 ...
## $ magnet_dumbbell_z       : num  -0.7272 -0.5841 -0.0333 0.0525 1.9051 ...
## $ roll_forearm            : num  0.992 0.695 0.899 -0.314 -1.945 ...
## $ pitch_forearm           : num  1.373 -1.008 -1.542 -0.382 -0.459 ...
## $ yaw_forearm             : num  1.322 0.838 0.712 -0.189 -0.652 ...
## $ max_picth_forearm       : num  0.621 0.578 0.518 -0.858 0.227 ...
## $ min_pitch_forearm       : num  0.796 1.149 0.874 0.522 -0.582 ...
## $ amplitude_roll_forearm  : num  0.8211 0.7235 0.1775 -0.9536 -0.0388 ...
## $ amplitude_pitch_forearm : num  -0.193 -0.484 -0.317 -0.945 0.581 ...
## $ var_accel_forearm       : num  -0.652 -0.442 -0.773 -0.858 0.163 ...
## $ gyros_forearm_x         : num  0.9251 1.5291 0.0349 1.9424 -1.4434 ...
## $ gyros_forearm_y         : num  -1.573 -1.314 -0.393 0.291 1.406 ...
## $ gyros_forearm_z         : num  -1.21 -0.529 0.234 2.757 1.097 ...
## $ magnet_forearm_y        : num  0.0704 0.802 0.6191 0.7863 -2.3014 ...
## $ magnet_forearm_z        : num  0.604 1.295 1.052 0.345 -0.815 ...

predanswers <- predict(model, predpmltest) 

```

## show the prediction results
```{r}
predanswers

## 1  2  3  4  5  6  7  8  9 10 11 12 13 14 15 16 17 18 19 20 
## B  A  B  A  A  E  D  B  A  A  B  C  B  A  E  E  A  B  B  B 
## Levels: A B C D E
```

